---
title: "Fundamentals"
author: "Kieran"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Fundamentals

## Vectors
* c - the combination function keyword
* A sequence of data elements, can be thought of as an array.
  + A single number considered a vector too.
  + Note all numbers are by default stored as double
* It is assumed you can only find data elements of the same type
  + Therefore a vector with different types is converted to 1 type
* Access elements in vectors using []
  + The index starts at 1
  
### Vector Operations
* All vector operations can be performed without the need of loops
  + Each element in the vector is compared to the other vector
  + In the case where a vector size differs
    + Vector Recycling will be used. 
      + The condition here is that the smaller vector needs to be a multiple of the bigger function
```{r}
N <- 100
a <- rnorm(N)
b <- rnorm(N)

c <- a * b
```
      
### Creating Vectors
```{r}
numVector <- c(3,5,10,25,50)
is.numeric(numVector)
is.integer(numVector)
is.double(numVector)

intVector <- c(1L, 2L, 5L)
is.integer(intVector)

charVector <- c("a", "B23", "Hello", 7)
charVector

#Access individual elements using []
charVector[1]
charVector[-3]
charVector[1:3] # Choose range of indexes
charVector[c(1,3,4)] # Choose specific indexes
```

#### seq
```{r}
seq(1,15) # Use sequence instead of : to adjust different params such as step
1:15

seq(1,15,2)
```

#### rep
```{r}
rep(3, 10) # Repeats an element
rep("c", 10)
rep(c(5,10), 10)
```